Legacy Data Confounds Genomics Studies

### 2019-03-27 PlottingScripts/ProcessPlotData.Rmd
update

#save to file
saveRDS(overLap,
            paste0(DataPath,
                   'overLap.rds'))
```
## Compare Models
```{r}

```

@@

#bcftools doesn't take rsID
#this deals with indels and other mulitallelic variants
vcftools \
--vcf $DataPath/Significant_Positions.vcf \
--snps $DataPath/Significant_rsID.txt \
--recode --recode-INFO-all \
--out $DataPath/Significant_Positions

@@

```

```{bash}
DataPath='/Users/luke/Documents/QualityPaper/ProcessedData'
RawDataPath='/Users/luke/genomes/genomes/Han90/90_Han_Chinese_SNP_gatkvqsr_20170424_chr'
VCF='_validator.vcf.gz'

@@

#bcftools doesn't take rsID
#this deals with indels and other mulitallelic variants
vcftools \
--vcf $DataPath/Significant_Positions_Han.vcf \
--snps $DataPath/Significant_rsID.txt \
--recode --recode-INFO-all \
--out $DataPath/Significant_Positions_Han

/Users/luke/bin/plink_mac/plink \
--vcf $DataPath/Significant_Positions_Han.recode.vcf \
--make-bed \
--out $DataPath/Significant_Positions_Han
```

##Full Model
```{r}
sharedID <- fread(paste0(DataPath,
                          '83samples.txt'), 
                  header = FALSE,
                  col.names = 'sample')

kgp_bim <- fread(paste0(DataPath,
                         'Significant_Positions.bim'), 
                  col.names = c('Chr','rsID','X',
                                'Pos','Ref','Alt')) %>%
  select(rsID)

kgp_fam <- fread(paste0(DataPath,
                         'Significant_Positions.fam'), 
                  col.names = c('familyID','sample',
                                'father','mother',
                                'sex','pheno')) %>%
  select(sample)

kgp_genotype <- read.bed(paste0(DataPath,'Significant_Positions'))
rownames(kgp_genotype) <- kgp_bim$rsID
colnames(kgp_genotype) <- kgp_fam$sample
#Only keep the 83 shared samples
keep <- colnames(kgp_genotype)[(colnames(kgp_genotype) 
                                 %in% sharedID$sample)]
kgp_genotype <- kgp_genotype[,keep]
kgp_genotype <- kgp_genotype[!duplicated(rownames(kgp_genotype)),]
kgp_genotype[1:5,1:5]

kgp1 <- melt(kgp_genotype, 
           varnames = c('rsID','sample'), 
           value.name = 'Genotype')
kgp <- 
		kgp1 %>% 
		group_by(rsID) %>% 
		summarize(kgp.AF = sum(Genotype))

kgp <- kgp[which(kgp$kgp.AF > 0),]
#read Han Genotypes
han_bim <- fread(paste0(DataPath,
                         'Significant_Positions_Han.bim'), 
                  col.names = c('Chr','rsID','X',
                                'Pos','Ref','Alt')) %>%
  select(rsID)

han_fam <- fread(paste0(DataPath,
                         'Significant_Positions_Han.fam'), 
                  col.names = c('familyID','sample',
                                'father','mother',
                                'sex','pheno')) %>%
  select(sample)

han_genotype <- read.bed(paste0(DataPath,'Significant_Positions_Han'))
rownames(han_genotype) <- han_bim$rsID
colnames(han_genotype) <- han_fam$sample
#Only keep the 83 shared samples
keep <- colnames(han_genotype)[(colnames(han_genotype) 
                                 %in% sharedID$sample)]
han_genotype <- han_genotype[,keep]
han_genotype <- han_genotype[!duplicated(rownames(han_genotype)),]
han_genotype[1:5,1:5]

han1 = melt(han_genotype, 
           varnames = c('rsID','sample'), 
           value.name = 'Genotype')

han <- 
		han1 %>% 
		group_by(rsID) %>% 
		summarize(han.AF = sum(Genotype))

#merge 1kGP and han resequenced data.
combo <- merge(han, kgp, by='rsID', all.y=T)
#missing sites replaced with 0
combo[is.na(combo)] <- 0

#save to file
saveRDS(combo,
            paste0(DataPath,
                   'HanKgp_AF.rds'))

combo
```

@@

>>>>

#save to file
saveRDS(overLap,
            paste0(DataPath,
                   'overLap.rds'))
```

@@

#bcftools doesn't take rsID
#this deals with indels and other mulitallelic variants
vcftools \
--vcf $DataPath/Significant_Positions.vcf \
--snps $DataPath/Significant_rsID.txt \
--keep $DataPath/83samples.txt \
--recode --recode-INFO-all \
--out $DataPath/Significant_Positions

@@

#Get Allele Frequencies
/Users/luke/bin/plink_mac/plink \
--bfile $DataPath/Significant_Positions \
--freq \
--out $DataPath/Significant_Positions
```

##Extract Significant Positions from Han Resequence Data
```{bash}
DataPath='/Users/luke/Documents/QualityPaper/ProcessedData'
RawDataPath='/Users/luke/genomes/genomes/Han90/90_Han_Chinese_SNP_gatkvqsr_20170424_chr'
VCF='_validator.vcf.gz'

@@

#bcftools doesn't take rsID
#this deals with indels and other mulitallelic variants
vcftools \
--vcf $DataPath/Significant_Positions_Han.vcf \
--snps $DataPath/Significant_rsID.txt \
--keep $DataPath/83samples.txt \
--recode --recode-INFO-all \
--out $DataPath/Significant_Positions_Han

/Users/luke/bin/plink_mac/plink \
--vcf $DataPath/Significant_Positions_Han.recode.vcf \
--make-bed \
--out $DataPath/Significant_Positions_Han

#Get Allele Frequencies
/Users/luke/bin/plink_mac/plink \
--bfile $DataPath/Significant_Positions_Han \
--freq \
--out $DataPath/Significant_Positions_Han
```

@@

##OMNI CHIP
```{r}
omni <- fread(paste0(DataPath,
                     'HumanOmni2-5-8-v1-2-A-b138-rsIDs.txt'),
              header = T)
omni <- merge(omni, 
              plt, 
              by.x='RsID', 
              by.y='rsID')
omni <- omni[which(omni$RsID != '.'),]

#save to file
saveRDS(omni,
            paste0(DataPath,
                   'omni.rds'))
```

### 2019-03-27 legacy.tex
update

Lists of suspicious variants and individuals with low $Q$ are provided as supplementary material \luke{link to file here} \sgcomment{Add this after all variants identified.}.

@@

This method identifies a total of 24,390 variants associated to $Q$ (Figures \ref{NRS_Manhattan},\ref{NRI_Manhattan}, \ref{RS_Manhattan}, and \ref{RI_Manhattan}). 
To account for the large number of tests, we used a two-stage Benjamini \& Hochberg step-up FDR-controlling procedure to adjust the p-values using a nominal Type-I error rate $\alpha = 0.01$ \citep{Benjamini2006}. We tested SNPs, INDELs and repetitive regions separately as they may have different error rates (Table \ref{sigTable}).

@@

\begin{figure}[h]
\centering
\includegraphics[width=15cm,keepaspectratio]{./Figures/dev_Pop_PC_PopPC.jpg}
\caption{Comparison of three logistic regression models for testing association to $Q$.
These methods model each genotype as a logistic function using principal components (PC), Population membership (Pop) or both (PopPC) as a covariate.
In these plots we are comparing the deviance from the null model in 50,000 variants from chromosome 22.}  
\label{CompareModel}
\end{figure}

@@

We did a similar analysis using the variants identified in the \textit{5PC} model present in the 83 resequenced individuals.
Of the 24,390 suspicious variants identified globally, only 7,607 of them are found in the 83 individuals in the original 1kGP data. (See Figure \ref{90HanSFS_full})
From this subset, only 613 of these variants are present in the resequenced data, the rest are missing entirely.
Figure \ref{90HanSFS_full} shows a weak correlation to allele frequency for the variants present in the resequenced data.
Whereas the vast majority of the SNPs are absent in resequenced data, a high proportion of INDELs are fixed in the resequenced data.
A number of variants shared across cohorts suggests that roughly 10\% of the variants identified are real but genotyped incorrectly.

@@

We found that 219 of the variants we identified as being associated to $Q$ were present on Illumina's Omni 2.5 chip. This does not mean that these variants are false positives, but rather, that in the 1kGP, individuals with low $Q$ were less likely to be called as having a variant in that position. Moreover, using the 1kGP as a reference panel might impact the population allele frequencies observed for these positions.

@@

\begin{table}[h]
\begin{tabular}{l  r r}
                      & {Repeat}  & {Non-Repeat}       \\ \hline
{SNP}  & 8,393,467 & 26,241,186   \\  
{INDEL} &  756,889  & 2,309,280 \\ \hline
\end{tabular}
\caption{Number of variants included in the analysis by category}
\label{totTable}
\end{table}

@@

@@

@@

\begin{figure}[h]
\centering
\includegraphics[width=12cm,keepaspectratio]{./Figures/Han_1kGP_SFS_FullModel.jpg}
\caption{Site frequency spectrum plot comparing the original 1kGP data to the high depth resequence data. This plot shows  11,946 significant SNPs and INDELs identified in the reverse GWAS model including all populations. Only 1,586 of the variants are present in the high depth resequenced individuals.}  
\label{90HanSFS_full}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\hsize,keepaspectratio]{./Figures/Hap.jpg}
\caption{Frequencies of $Q$-associated variants polymorphic in both 1000G and either HapMap or GnomAD. \sgcomment{Change title, check targeted  sequencing region. Specify proportion of SNPs not included} The blue and green dots in the HapMap plot are the variants reported in \citep{Mandage2017} and \citep{Kraja2011} respectively. There are 17 such variants in HapMap and 186 in the GnomAD.}
\label{HapMap_GnomAD}
\end{figure}

>>>>

@@

This method identifies a total of 24,390 variants associated to $Q$ (Figures \ref{NRS_Manhattan},\ref{NRI_Manhattan}, \ref{RS_Manhattan}, and \ref{RI_Manhattan}). 
To account for the large number of tests, we used a two-stage Benjamini \& Hochberg step-up FDR-controlling procedure to adjust the p-values using a nominal Type-I error rate $\alpha = 0.01$ \citep{Benjamini2006}. We tested SNPs, INDELs and repetitive regions separately as they may have different error rates (Table \ref{sigTable}).
Lists of suspicious variants and individuals with low $Q$ are provided as supplementary material \luke{link to file here}.

@@

@@

We did a similar analysis using the variants identified in the \textit{5PC} model present in the 83 resequenced individuals.
Of the 24,390 suspicious variants identified globally, only 12,086 of them are found in the 83 individuals in the original 1kGP data. (See Figure \ref{90HanSFS_full})
From this subset, only 1,637 of these variants are present in the resequenced data, the rest are missing entirely.
Figure \ref{90HanSFS_full} shows a relatively weak correlation to allele frequency for the variants present in the resequenced data.
A number of variants shared across cohorts suggests that roughly 10\% of the variants identified are real but genotyped incorrectly.

@@

We found that 613 of the variants we identified as being associated to $Q$ were present on Illumina's Omni 2.5 chip. This does not mean that these variants are false positives, but rather, that in the 1kGP, individuals with low $Q$ were less likely to be called as having a variant in that position. Moreover, using the 1kGP as a reference panel might impact the population allele frequencies observed for these positions.

@@

\begin{table}[h]
\begin{tabular}{l  r r}
                      & {Repeat}  & {Non-Repeat}       \\ \hline
{SNP}  & 6,312,620 & 19,846,786   \\  
{INDEL} &  586,342  & 1,770,315 \\ \hline
\end{tabular}
\caption{Number of variants included in the analysis by category}
\label{totTable}
\end{table}

@@

\begin{figure}[h]
\centering
\includegraphics[width=15cm,keepaspectratio]{./Figures/fits_Significant_Positions_CompareModels.jpg}
\caption{Comparison of three logistic regression models for testing association to $Q$.
These methods model each genotype as a logistic function using principal components (PC), Population membership (Pop) or LFA as an offset.
In these plots we are comparing the deviance from the null model in the 24,390 variants identified using the LFA model.}  
\label{CompareModel}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=12cm,keepaspectratio]{./Figures/Han_1kGP_SFS_FullModel.jpg}
\caption{Site frequency spectrum plot comparing the original 1kGP data to the high depth resequenced data. 
This plot shows  12,086 significant variants identified in the reverse GWAS model including all populations.
 Only 1,637 of the variants are present in the high depth resequenced individuals.}  
\label{90HanSFS_full}
\end{figure}

\begin{figure}[h]
\includegraphics[width=\hsize,keepaspectratio]{./Figures/Omni_AF.jpg}
\caption{We found that 613 of the variants we identified as being associated to $Q$ were present on Illumina's Omni 2.5 chip. 
This does not mean that these variants are false positives, but rather, that in the 1kGP, individuals with low $Q$ were less likely to be called as having a variant in that position. 
Moreover, using the 1kGP as a reference panel might impact the population allele frequencies observed for these positions.}
\label{HapMap_GnomAD}
\end{figure}

@@
